---
title: "Homework 4"
author: "Your Name Here"
subtitle: "Spring 2023"
header-includes:
    - \usepackage{amsmath}
    - \usepackage{amsthm}
output: pdf_document
---

```{r setup, include=FALSE}
#Use this code chunk to include libraries, and set global options.

knitr::opts_chunk$set(echo = TRUE)

```

### Instructions

- This homework is due in Gradescope on Wednesday May 3 by midnight PST.

- Please answer the following questions in the order in which they are posed. 

- Don't forget to knit the document frequently to make sure there are no compilation errors. 

- When you are done, download the PDF file as instructed in section and submit it in Gradescope. 

* * *

### Exercises

1. (MOM vs.MLE) Suppose $X_1,X_2,\dots,X_n$ are independently drawn from the PDF
$$f(x) = (\theta_0 +1)x^{\theta_0} \ \ \ 0 \leq x \leq 1$$
where the parameter $\theta_0 > -1$ so the PDF is non-zero and integrates to 1.

a. Derive $\widehat{\theta}_0^{mom}$, the method of moments estimator of $\theta_0$.

b. Derive $\widehat{\theta}_0^{mle}$, the maximum likelihood estimator of $\theta_0$. Be sure to show
   - likelihood function
   - log-likelihood function
   - first derivative condition
   - second derivative test
   
c. Find $\widehat{\theta}_0^{mom}$ based on the sample below.

```{r echo =T}
x <- c(0.90, 0.78, 0.93, 0.64, 0.45, 0.85, 0.75, 0.93, 0.98, 0.78)

```

d. Make a plot of the log-likelihood function for the data from part c.and calculate $\widehat{\theta}_0^{mle}$.

2. (Light bulbs) A set of	cheap light	bulbs have a lifetime (in hours) which is	exponentially distributed with unknown rate	$\lambda_0$:
$$f(x) = \lambda_0 exp(-\lambda_0\:x), \ \ 0 < x $$

   Choosing	a random	sample of ten light	bulbs, they are turned on simultaneously and observed for 48 hours. During this period, six bulbs went out, at times $x_1,x_2,\dots,x_6$. At the end of the experiment, four light bulbs were still working. 

  a.  Derive the likelihood function $L(\lambda)$.   (Hint:  we can model this as observing values for $X_1,X_2,\dots,X_6$ which are exponential random variables and $Y_1,Y_2,Y_3,Y_4$ which are Bernoulli random variables which are 1 or 0 depending on whether the lifetime $X$ is larger than 48 or not. The likelihood function is the product of the six exponential density functions and the four Bernoulli PMF.)

  b. Derive an expression for the MLE of $\lambda_0$ showing your work. Verify it is the global maximum of the likelihood function. 
  
3. (SLR through origin) Suppose $X_1,X_2,\dots,X_n \overset{i.i.d.}{\sim} Norm(a_i \mu_0, 1)$ where the $a_i$ are known constants. (FYI: This is a Simple Linear Regression (SLR) model which is forced to go through the origin since there is no intercept)

a. Write the likelihood function $L(\mu)$ and also the log-likelihood function $\ell(\mu)$.

b. Derive an expression for $\widehat{\mu}_0^{mle}$, the MLE of $\mu_0$.  (Please show your steps clearly, including the second derivative test)

c. Is $\widehat{\mu}_0^{mle}$ an unbiased estimator of $\mu_0$? Show your work.

d. Derive the standard error of $\widehat{\mu}_0^{mle}$. 

4. (Two scientists) A	scientist	has	obtained	two	random	samples:	one	of	size	$n_1$	from	an	exponential	
distribution	with	mean	$\theta_0$ and	another	of	size	$n_2$	from	an	exponential	
distribution	with	mean	$k \theta_0$,	where	k	is	a	known	number,	but	$\theta_0$ is	unknown.	
The	scientist	has	computed	the	MLEs for $\theta_0$ - let's call them $\widehat{\theta}_{0}^{mle1}$ and $\widehat{\theta}_0^{mle2}$ from each	of	the	samples.	Now	they
want	a	single	estimate	of	$\theta_0$,	so	they	ask	two	statisticians	for	advice.	One	suggests	
finding	the	linear	combination	 $a \widehat{\theta}_0^{mle1} + (1-a) \widehat{\theta}_0^{mle2}$,	with	the	smallest	variance.	
The	other	suggests	finding	the	MLE	from	the	combined	sample.	Show	that	both	
methods	yield	the	same	answer.	

   To help us with the grading, please 
- clearly separate the work pertaining to derivation of $\widehat{\theta}_0^{mle1}$ and $\widehat{\theta}_0^{mle2}$
- clearly show your steps (for example, for finding $a$)
- clearly highlight your final estimators in each case by stating them. 