---
title: "Problem Section 1"
subtitle:  "Robustness of the t-test"
graphics: yes
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


### Exercises 

1. Every t test makes the same explicit assumption - namely, that the set of $n$ data points - $X_1,X_2, \dots, X_n$ - are normally distributed. 
If the normality assumption is not satisfied, then the ratio 
$$T = \frac{\bar{X} - \mu_0}{\frac{S}{\sqrt{n}}}$$ will not have a t-distribution. However, whether or not the validity of the t-test is compromised depends on how different the actual distribution of the statistic $T$ is from the $t$ distribution.

In this exercise, you will investigate the sensitivity of the $T$ ratio to violations of the normality assumption by
simulating samples of size $n$ from selected distributions and comparing the resulting
histogram to the $t$ distribution with $n-1$ degrees of freedom.


a. Simulate $B = 10,000$ samples of size $n = 6, 15$ each from a Unif(0,1) distribution. For each sample, calculate $$t = \frac{\sqrt{n}(\bar{x} - \mu_0)}{s}$$ where $\mu_0 = \frac{1}{2}$ is the mean of the uniform distribution. Create a histogram of the $t$ ratios and superimpose the $t$ distribution with $n-1$ degrees of freedom. What do you notice?

```{r unif-dist}
set.seed(2737)

```

b. Repeat part a. for samples drawn from an exponential distribution with rate $\lambda_0 = 2$. (Note: $\mu_0 = 1/2$ for this distribution also) What do you notice?

```{r exp-dist}
set.seed(88)
```


2. Your simulations in problem 1 should show that the distribution of $$T = \frac{\sqrt{n}(\bar{X} - \mu_0)}{S}$$ will become increasingly similar to a $t_{n-1}$ distribution as $n$ increases, regardless of which distribution you sample from. Can you explain why this happens?

3. What is $Cov(X, X)$?

4. Two draws are made at random from the box below
\begin{center}
$\fbox{ \fbox{1} \fbox{2} \fbox{3}  }$.
\end{center}

Let $X$ denote the number on the first randomly selected ticket and $Y$ the second. The joint PMF of $\langle X, Y \rangle$ is shown below.

\begin{table}[h]
\centering
\begin{tabular}{|cccccc|ccccc|}
& & \multicolumn{3}{c}{With replacement} & & & \multicolumn{3}{c}{Without replacement} & \\ \hline
& & \multicolumn{3}{c}{$X$} & $f_{Y}$ & & \multicolumn{3}{c}{$X$} & $f_{Y}$ \\ 
& & 1 & 2 & 3 & & & 1 & 2 & 3  & \\ \hline
& 1 & 1/9 & 1/9 & 1/9 & 1/3 & & 0 & 1/6 & 1/6  & 1/3 \\ 
$Y$ & 2 & 1/9 & 1/9 & 1/9 &  1/3 &  & 1/6 & 0 & 1/6 &  1/3 \\ 
& 3 & 1/9 & 1/9 & 1/9 & 1/3 &   & 1/6 & 1/6 & 0& 1/3 \\ \hline
& $f_{X}$ & 1/3 & 1/3 & 1/3 & 1.00  & & 1/3 & 1/3 & 1/3  & 1.00
\end{tabular}
\end{table}


a. When the draws are made with replacement, $Cov\left[X, Y \right]=0$. Why?

b.  Find $Cov\left[ X, Y \right]$ when the draws are made without replacement. Does the sign make sense?