---
title: "Problem Section 3"
subtitle:  "Two sample inference for means"
graphics: yes
output: pdf_document
header-includes: 
- \usepackage{amssymb, amsmath, amsfonts}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```



### Exercises 


1. A group of college students interested in the
effect of stepping exercises on the heart conducted an experiment in which subjects
were randomly assigned to a stepping exercise on either a low step (coded "low") or a high step (coded "high"). Each subject started with a resting heart rate and performed the exercise for 3 minutes, at which time his or her exercise heart rate was measured.
The data are in the file `exercise.csv`. 

a. The students first want to see if the random assignment of subjects
to the groups did a satisfactory job of equalizing the mean resting heart rates between the groups. Explain how to use a confidence interval to look for evidence of a problem with the random assignment. Then go ahead and implement it.

b. Next, you want to test whether there is a difference in heart rates between the two groups. Explain how to use a confidence interval to examine this question. Then go ahead and implement it.

2. For comparing two means using the t-test, we need to assume the data are normally distributed. With small $n$, it is difficult to judge the shape of a population distribution. Also, if the distribution is skewed the mean is less relevant, but we may still want to test the hypothesis that the two groups have the same distribution. A \emph{permutation test} is a non-parametric method for testing the null hypothesis that the population distributions are identical without specifying their shape.

Consider the following data to determine whether dogs prefer vocal praise or petting. The response variable is the time, in seconds, the animal spent interacting with its owner.

```{r }
dogs <- tibble( 
  group = rep(c("pet","praise"), times = c(7,7)),
  times = c(114, 203, 217,254, 256, 284, 296, 4,7, 24, 25, 48, 71, 294)
  )
```

a. Make a violin plot to compare the distributions. 

b. As the true distributions are potentially highly skewed, we focus on the medians. Calculate the observed difference in the medians (pet - praise). This is the observed value of our test statistic. Save it in a variable called `obs_diff_in_median`.

c. We test the null hypothesis of identical population distributions against the alternative that the population median is higher for petting using a \textbf{permutation test}. The key to permutation testing is this: 

\emph{if $H_0$ is true, meaning the population distributions are the same, then if we pool all the data and "reshuffle" the pooled values into two new groups of size $n=7$ and $m=7$, we will be sampling from the same (null) distribution. }


The following code implements this method once. We set a seed any time random number generation is involved. After dividing the pooled data into two new groups, we re-calculate the difference in the medians.

```{r label="shuffle-data"}
set.seed(1414)


#sample 7 row numbers
i <- sample(1:nrow(dogs), size = 7, replace=FALSE)    


new_sample1 <- dogs[i,]$times
new_sample2 <- dogs[-i,]$times

median(new_sample1)-median(new_sample2)
  
                                
```

d. We can generate the sampling distribution of the test statistic by generating \emph{all} possible ways to form the two samples of size $n$ and $m$ from the pooled data,  and compute the difference in median for each. Alternately, we could select a subset of the samples randomly and calculate the difference in medians for those. This is the approach we will take here.

The P-value is for the permutation test is the fraction of times the difference between the sample medians is at least `obs_diff_median`.  


Write code to create B=10,000 new "reshuffles" of the pooled data. For each reshuffle, calculate the difference in sample medians. Then make a plot to visualize the null sampling distribution of this statistic. Finally, calculate the  P-value for the permutation test.

```{r}
set.seed(24)
B<-10000

```